An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-09-23 latex_files/ch2/2_1_implicit_models.tex
multiple edits until Sec 2.4

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic
of these models is that we can sample (simulate) data points,
but that we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.

>>>>

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample (simulate) data points, but that we cannot evaluate
the likelihood. Formally, a simulator-based model is described as a
parameterised family of probability density functions $\{
p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
data generating machine $\simulator$\footnote{The subscript $r$ in $M_r$
indicates the \textit{random} simulator. In the next chapters we will
introduce $M_d$ witch stands for the \textit{deterministic}
simulator.} that produces samples $\yb$ in a stochastic manner for any
given a set of parameters $\thetab$.% i.e.\
$M_r(\thetab) \rightarrow \yb$.

### 2021-09-23 latex_files/ch2/2_2_LFI_background.tex
multiple edits until Sec 2.4

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample, or simulate, data points, but that we cannot
evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is a parameterised family
of probability density functions $\{ p(\yb|\thetab) \}_{\thetab}$,
whose closed-form is either unknown or computationally
intractable. Whereas evaluating $p(\yb|\thetab)$ is intractable,
sampling is feasible. Practically, a simulator can be understood as a
random black-box machine (computer code) $M_r$ that generates samples
$\yb$ in a stochastic manner for any given a set of parameters
$\thetab$. We can isolate the randomness by introducing the stochastic
nuisance variables $\vb \sim p(\vb)$ so that the simulator becomes a
deterministic mapping $M_d$ that maps $(\thetab, \vb)$ to the data
$\yb$. The distribution $p(\vb)$ is determined by the random draws
performed when running $M_r$. Importantly, we can fix the value of
$\vb$ while changing $\thetab$ by fixing the random seed used with
$M_r$.

@@

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. The main
difficulty is that the likelihood function $L(\thetab) =
p(\data|\thetab)$ is intractable. To better see the sources of the
intractability, and to address them, we go back to the basic
characterisation of the likelihood as the (rescaled) probability for a
parameter of the model to generate data $\yb$ that is similar to the
observed data $\data$. More formally, the likelihood $L(\thetab)$
equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in \region(\data))
\end{equation}
where $c_\epsilon$ is a proportionality factor that depends on
$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
that is defined via a distance function $d$, i.e.\ $\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}$.

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability $\Pr(M_r(\thetab) \in \region(\data))$, the second
is the limit of $\epsilon \to 0$.  Whilst the probability can be
approximated with samples, this becomes computationally infeasible if
$\epsilon$ is too small. Hence, a large class of inference methods
work with $\epsilon >0$, which leads to the approximate likelihood
function $L_{d, \epsilon}(\thetab)$,
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)), \quad \text{where  } \epsilon > 0.
\end{equation}
Most often in likelihood-free inference, we are interested in the posterior distribution of the parameters $\thetab$ given some observed data. The approximate likelihood function in \eqref{eq:approx_likelihood} implies the approximate posterior  $p_{d,\epsilon}(\thetab|\data)$, 
\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

>>>>

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample, or simulate, data points, but that we cannot
evaluate the likelihood. Formally, a simulator-based model is a
parameterised family of probability density functions $\{
p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either unknown or
computationally intractable. Whereas evaluating $p(\yb|\thetab)$ is
intractable, sampling is feasible. Practically, a simulator can be
understood as a random black-box machine (computer code) that
generates samples $\yb$ in a stochastic manner for any given a set of
parameters $\thetab$. We can isolate the randomness by introducing the
stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator
becomes a deterministic mapping $\simulator$ that maps
$(\thetab, \vb)$ to the data $\yb$, i.e.\
$\yb=\simulator(\thetab,\vb)$. The distribution $p(\vb)$ is determined
by the random draws performed when running the computer
code. Importantly, we can fix the value of $\vb$ while changing
$\thetab$ by fixing the random seed used.

@@

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. Denoting observed
data by $\data$, the main difficulty is that the likelihood function
$L(\thetab) = p(\data|\thetab)$ is intractable. To better see the
sources of the intractability, and to address them, we go back to the
basic characterisation of the likelihood as the (rescaled) probability
for a parameter of the model to generate data $\yb$ that is similar to
the observed data $\data$. More formally, the likelihood $L(\thetab)$
equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data))
\end{equation}
where $c_\epsilon$ is a proportionality factor that depends on
$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
that is defined via a distance function $d$, i.e.\ $\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}$.

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$, the second
is the limit of $\epsilon \to 0$.  Whilst the probability can be
approximated with samples, this becomes computationally infeasible if
$\epsilon$ is too small. Hence, a large class of inference methods
work with $\epsilon >0$, which leads to the approximate likelihood
function $L_{d, \epsilon}(\thetab)$,
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)), \quad \text{where  } \epsilon > 0.
\end{equation}
Most often in likelihood-free inference, we are interested in the posterior distribution of the parameters $\thetab$ given some observed data. The approximate likelihood function in \eqref{eq:approx_likelihood} implies the approximate posterior  $p_{d,\epsilon}(\thetab|\data)$, 
\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

### 2021-09-23 latex_files/ch2/2_3_OMC.tex
small changes
multiple edits until Sec 2.4

Our description of OMC \citep{Meeds2015}
follows \citet{Ikonomov2019} who base their explanation of OMC on the
approximate likelihood function in \eqref{eq:approx_likelihood}. With
the indicator function
%
\begin{equation} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise} 
    \end{array} \right. \end{equation}
%
we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
  \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
\end{gather}
which we can approximate as a sample average
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
  \text{ where } \yb_i = M_d(\thetab, \vb_i), \quad \vb_i \sim p(\vb). \label{eq:alt_view}
\end{equation}
We thus approximate the likelihood of a specific $\thetab$ by counting
the fraction of samples that lie inside the area around the
observations. The fact that we isolated the randomness of the
simulator via the $\vb_i$ has two crucial consequences: First, we can
sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
different values of $\thetab$. Secondly, since
$\yb_i=M_d(\thetab, \vb_i)$ is a function of $\thetab$, checking
whether $\yb_i$ is contained in region $\region (\data)$ for a
particular value of $\thetab$ is the same as checking whether
$\thetab$ is in the acceptance region $\mathcal{S}_i = \{ \thetab:
M_d(\thetab, \vb_i) \in \region(\data) \}$. This leads to the
likelihood approximation
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\mathcal{S}_i}(\thetab) \label{eq:alt_view_theta}
\end{equation}
and the corresponding approximate posterior
\begin{equation} \label{eq:approx_posterior_omc}
  p_{d,\epsilon}(\thetab|\data) \propto
  p(\thetab) \sum_i^N  \indicator{\mathcal{S}_i}(\thetab).
\end{equation}
As argued by \citet{Ikonomov2019}, this derivations provide a unique
perspective for likelihood-free inference by shifting the focus onto
the geometry of the acceptance regions $\mathcal{S}_i = \{ \thetab:
M_d(\thetab, \vb_i) \in \region(\data)\}$. Indeed, the task of
approximating the likelihood and/or posterior becomes a task of
characterising the sets $\mathcal{S}_i$.

OMC by \citet{Meeds2015} assumes that the distance $d$ is the
Euclidean distance between summary statistics $\Phi$ of the observed
and generated data, and that the $\mathcal{S}_i$ can be well
approximated by infinitesimally small ellipses. These assumptions
leads to an approximation of the posterior in terms of weighted
samples $\thetab_i^*$. In more detail, OMC considers the limit of
$\epsilon \to 0$ and represents the sets $\mathcal{S}_i$ by points
$\thetab_i^*$ which are those parameter configurations that achieve
the smallest distance between observed and simulated data for each
realisation $\vb_i \sim p(\vb)$, i.e.\ they are solutions to the
deterministic optimisation problems
\begin{equation}
\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
\end{equation}
The weighting for each $\thetab_i^*$ is proportional to the prior
density at $\thetab_i^*$ and inversely proportional to the determinant
of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
further details on OMC we refer the reader to \citep{Meeds2015,
Ikonomov2019}.

>>>>

Our description of OMC \citep{Meeds2015}
follows \citet{Ikonomov2019} who base their explanation of OMC on the
approximate likelihood function in \eqref{eq:approx_likelihood}. With
the indicator function
%
\begin{equation} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise} 
    \end{array} \right. \end{equation}
%
we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
  \int_{\yb \in B_{d,\epsilon}(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
\end{gather}
which we can approximate as a sample average
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
 \quad \text{ where } \yb_i = \simulator(\thetab, \vb_i),\; \vb_i \sim p(\vb). \label{eq:alt_view}
\end{equation}
We thus approximate the likelihood of a specific $\thetab$ by counting
the fraction of samples that lie inside the area around the
observations. The fact that we isolated the randomness of the
simulator via the $\vb_i$ has two crucial consequences: First, we can
sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
different values of $\thetab$. Secondly, since
$\yb_i=\simulator(\thetab, \vb_i)$ is a function of $\thetab$,
checking whether $\yb_i$ is contained in region $\region (\data)$ for
a particular value of $\thetab$ is the same as checking whether
$\thetab$ is in the acceptance region $\accregioni
= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}$. This
leads to the likelihood approximation
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab) \label{eq:alt_view_theta}
\end{equation}
and the corresponding approximate posterior
\begin{equation} \label{eq:approx_posterior_omc}
  p_{d,\epsilon}(\thetab|\data) \propto
  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
\end{equation}
As argued by \citet{Ikonomov2019}, these derivations provide a unique
perspective for likelihood-free inference by shifting the focus onto
the geometry of the acceptance regions $\accregioni$. Indeed, the task
of approximating the likelihood and/or posterior becomes a task of
characterising the sets $\accregioni$.

OMC by \citet{Meeds2015} assumes that the distance $d$ is the
Euclidean distance between summary statistics $\Phi$ of the observed
and generated data, and that the $\accregioni$ can be well
approximated by infinitesimally small ellipses. These assumptions leads to an approximation of the posterior in
terms of weighted samples $\thetab_i^*$ that achieve the smallest
distance between observed and simulated data for each realisation
$\vb_i \sim p(\vb)$, i.e.\
\begin{equation} \label{eq:omc_opt_prob}
\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
\end{equation}
The weighting for each $\thetab_i^*$ is proportional to the prior
density at $\thetab_i^*$ and inversely proportional to the determinant
of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
further details on OMC we refer the reader to \citep{Meeds2015,
Ikonomov2019}.

### 2021-09-23 latex_files/ch2/2_4_ROMC.tex
multiple edits until Sec 2.4

The simplifications introduced by OMC, although beneficial from a
computational point-of-view, they suffer from some drawbacks:

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$
  shrinks to a single point $\thetab_i^*$. This simplification adds
  significant approximation error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. Using only local information for estimating the
  weight can also be misleading. For example, if $g_i$ is almost flat
  at $\thetab_i^*$, i.e.
  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then w_i \rightarrow
  \infty$, dominating the posterior.
\item There is no way to solve the optimisation problem
  of~\eqref{eq:omc_opt_prob} when
  $||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2$ is not differentiable.
\end{itemize}

\subsubsection{Sampling and computing expectation in ROMC}

ROMC resolves the failure-modes of OMC, mainly, by approaching the
acceptance regions $\mathcal{S}_i$ in a more sophisticated way;
instead of collapsing the regions into single points, it approximates
them with a bounding box. Afterwards, a uniform distribution defined
on the bounding box area, is used as the \emph{proposal distribution}
for importance sampling. Defining as $q_i$ the uniform distribution
defined on the $i-th$ bounding box, weighted sampling is performed by
drawing samples from $q_i$ and assigning the weights as
in~\eqref{eq:sampling}.

\begin{gather}
  \thetab_{ij} \sim q_i \\  \label{eq:sampling}
  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
\end{gather}

\noindent

Having defined the procedure for obtaining weighted samples, any
expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
as,

\begin{equation} \label{eq:expectation}
  E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}

The constrution of the proposal regions is performed in two steps; (a)
solving the optimisation problems as in OMC and (b) constructing the
bounding box regions.

\subsubsection*{Solving the deterministic optimisation problems}

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
we search for a point $\thetab^*_i$ such that
$d(M_d(\thetab^*_i,\vb_i), \data) < \epsilon$. For notation
convenience, we name the distance $d(M_d(\thetab,\vb_i), \data)$ as
$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
optimisation problem:

\begin{subequations}
\begin{alignat}{2}
  &\!\min_{\thetab}        &\qquad& d_i(\thetab) = d(M_d(\thetab^*,\vb_i), \data)\label{eq:optProb}\\
  &\text{subject to} & & d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) < \epsilon$. The nature of the generative
model $M_r(\thetab)$, specifies the properties of the objective
function $d_i$. If $d_i$ is differentiable any gradient-based
optimizer can be used for~\ref{eq:optProb}. The gradients
$\nabla_{\thetab} d_i$ can be either provided in closed form or
approximated by finite differences. In case $d_i$ is not
differentiable, Bayesian Optimisation~\cite{Shahriari2016} provides an
alternative choice. In this scenario, apart from obtaining an optimal
$\thetab_i^* $, a surrogate model $\hat{d}_i$ of the distance $d_i$ is
also fitted; $\hat{d}_i$ can then substitute the actual model in the
following steps\footnote{Such as proposal region construction, sampling etc}.

\subsubsection*{Proposal region estimation}

The acceptance region $\mathcal{S}_i$ is approximated with a bounding
box $\mathcal{\hat{S}}_i$ around it. The bounding box must contain the
part of the acceptance region \emph{around} $\thetab_i^*$; formally
$\{ \thetab : \thetab \in \mathcal{S}_i \quad \text{and} \quad
||\thetab - \thetab_i^*||_2 < M \}$. The condition
$||\thetab - \thetab_i^*||_2 < M$ restricts that if $\mathcal{S}_i$
contains many \emph{disjoint} sets of $\thetab$ that they fulfill
$d_i(\thetab) < \epsilon$, we want the bounding box to match only the
set that contains $\thetab_i^*$. Ideally, we want the bounding box to
be as tight as possible to $\mathcal{S}_i$, to ensure high acceptance
rate in the importance sampling, and big enough for not discarding
valid parts of $\mathcal{S}_i$.

The bounding boxes are built in two steps; firstly, we define their
axes $\mathbf{v}_d$ in the $d$-dimensional space and, then, the limits
along the axes are obtained with a line-search method looking for the
smallest
$\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite
  direction along the search
  line}. Algorithm~\ref{alg:region_construction} describes the method
in-depth. After the bounding boxes construction, a uniform
distribution $q_i$ is defined on each bounding box, and is used as the
proposal region for importance sampling.

\subsubsection*{Fitting a local surrogate model $\hat{g}_i$}

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in $d_i(\thetab)$ outside
$\mathcal{\hat{S}}_i$. In the inference steps
(i.e. sampling~\eqref{eq:sampling}, computing
expectation~\eqref{eq:expectation}, evaluating the
posterior~\eqref{eq:approx_posterior_omc}), we need to evaluate the
distance function $d_i$ for checking if the samples obtained from the
uniform distribution $\thetab_{ij} \sim q_i$ lie inside the acceptance
region $\mathcal{S}_i$. Evaluating $d_i$ involves running the
deterministic simulator $M_d(\thetab_{ij}, \vb_i)$, which can be
inneficient when the model is big.\footnote{If Bayesian Optimization
  has been used for obtaining $\thetab_i^*$, then $\hat{d}_i$ can be
  used as an alternative.}. Hence, a \emph{light} surrogate model
$\Tilde{d}_i$ can be fit for approximating $d_i$ around $\theta_i^*$.

ROMC proposes fitting a simple quadratic model\footnote{Any ML
  regression model may be chosen as local surrogate. The choice should
  consider the properties of the local region (i.e. size,
  smoothness). } by sampling training points from the bounding box and
labeling them by evaluating $d_i$. This additional step adds an extra
effort in the training part, but promises a significant speed-up at
the inference phase. Such time-badget imbalance between the training
and the inference part is demanded in many ML contexts; it is usual to
be quite generous with the execution time at the training phase but
quite eager at the inference phase. Fitting a local surrogate model
aligns with this requirement.

>>>>

\citet{Ikonomov2019} showed that considering infinitesimally small ellipses can lead
to highly overconfident posteriors. We refer the reader to their paper
for the technical details and conditions for this issue to
occur. Intuitively, it happens because the weights in OMC are only
computed from information at $\thetab_i^*$. And using only local
information can be misleading. For example if the curvature of
$||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at $\thetab_i^*$
is nearly flat, the curvature alone may wrongly indicate that
$\accregioni$ is much larger than it actually is. We implement in our
software package the robust generalisation of OMC
by \citet{Ikonomov2019} that resolves this issue.

ROMC approximates the acceptance regions $\accregioni$ and defines
local proposal distributions $q_i(\thetab)$ on them from which the
posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
are assigned (importance) weights $w_{ij}$ that compensate for using
the proposal distributions $q_i(\thetab)$ and not the prior
$p(\thetab)$,
\begin{equation}
  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
  \label{eq:sampling}
\end{equation}
Given the weighted samples, any expectation
$\Ex_{p(\thetab|\data)}[h(\thetab)]$ of some function $h(\thetab)$, can be approximated as
\begin{equation} \label{eq:expectation}
  \Ex_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}
\citet{Ikonomov2019} considered uniform distributions as proposal
distributions so that the main work is to approximate the acceptance
regions $\accregioni$ and to represent them such that uniform sampling is
easy. The approximation of the acceptance regions contains two compulsory
and one optional step: (1) solving the optimisation problems as in OMC, (2)
constructing bounding boxes around $\accregioni$ and optionally, (3)
refining the approximation via a probabilistic surrogate model of the distance.
\subsubsection*{Solving the deterministic optimisation problems}
For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$, we
search for a point $\thetab^*_i$ such that
$d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon$. For notational
convenience, we denote the distance $d(\simulator(\thetab,\vb_i), \data)$ by
$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
optimisation problem:
\begin{subequations}
\begin{alignat}{2}      
  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) \le \epsilon$. If $d_i(\thetab)$ is
differentiable any gradient-based optimizer can be used
for~\ref{eq:optProb}. The gradients $\nabla_{\thetab} d_i(\thetab)$
can be either provided in closed form or approximated by finite
differences. In case $d_i$ is not differentiable, Bayesian
Optimisation~\cite{Shahriari2016} provides an alternative choice. In
this scenario, apart from obtaining an optimal $\thetab_i^* $, a
surrogate model $\hat{d}_i(\thetab)$ of the distance function
$d_i(\thetab)$ is also automatically obtained; $\hat{d}_i$ can then
substitute the actual distance function in downstream steps of the
algorithms, which can lead to computational gains if running the
simulator and hence evaluating the distance function $d_i(\thetab)$ is
computationally expensive.

\subsubsection*{Estimating the acceptance regions}
The acceptance region $\accregioni$ is approximated by a bounding box
$\accregionihat$ around it. Ideally, we want the bounding box to be as
tight as possible to $\mathcal{S}_i$, to ensure high acceptance rate
in the importance sampling, and big enough for not discarding valid
parts of $\mathcal{S}_i$. \TODO{did remove the issue of disconnected
sets; which needed to be handled by multiple restarts with the same
seed. Is this something that the code would support? It's expensive
because we don't know when we need it! I think it's an open research
question, not related to this software paper, so I omitted it.} The
bounding boxes are built in two steps. First, we define their axes
$\mathbf{v}_d$ based on the (estimated) curvature of the distance
$\thetab_i^*$, second, we determine the size of the box via a
one-dimensional line-search method along each axix, see
Algorithm~\ref{alg:region_construction} for the details. After the
bounding boxes construction, a uniform distribution $q_i$ is defined
on each bounding box, and is used as the proposal region for
importance sampling.

\subsubsection*{Refining the estimate via a local surrogate model (optional)}
When computing the weight $w_{ij}$ in \eqref{eq:sampling}, we need to
check whether the samples $\thetab_{ij} \sim q_i$ lie inside the
acceptance region $\accregioni$. This can be considered to be a
safety-mechanism that corrects for any inaccuracies in the
construction of $\accregionihat$ above. However, this check involves
evaluting the distance function $d_i(\thetab_{ij})$ and hence running
the simulator $\simulator(\thetab_{ij}, \vb_i)$, which can be
expensive if the model is complex.

\citet{Ikonomov2019} thus proposed to construct a surrogate model $\hat{d}_i(\thetab)$
of the distance function $d_i(\thetab)$ valid in $\accregionihat$ and
that is is cheap to evaluate. They used a simple quadratic model
whilst other regression models are, in principle, possible too. The
advantage of using a quadratic model is that it has ellipsoidal
isocontours, which thus naturally allowed \citet{Ikonomov2019} to
replace the bounding box approximation of $\accregioni$ with a
tigher-fitting ellipsoidal approximation.\footnote{The difference to
the infinitesimal ellipsoidal model in OMC is the estimation procedure: OMC uses
information at $\thetab_i^*$ whilst, here, information in
$\accregionihat$ is used, which results in a more stable fit.}

The training data for fitting the quadratic model is obtained by
sampling $\thetab \sim q_i$ and evaluating the corresponding distances
$d_i(\thetab)$. The generation of the training data adds an extra
computational cost, but leads to a significant speed-up when
evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
eliminated if Bayesian Optimization with a Gaussian processe (GP)
surrogate model was used to obtain $\thetab_i^*$ in the first place
(see above). In this case, we can use the GP model of the distance
function instead of $d_i(\thetab)$ to generate the training data. This
essentially replaces the global GP model with a simpler local
quadratic model which is typically more robust.
